---
title: "Exploratory Data Analysis (EDA) Paper"
author: "Anthony Hu, Rishabh Sud, Aysha Ahmed"
date: '06/16/2021'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Data Import and Join

```{r}
library(tidyverse)
```

# Data
[Chicago Census](https://datahub.cmap.illinois.gov/dataset/1d2dd970-f0a6-4736-96a1-3caeb431f5e4/resource/8c4e096e-c90c-4bef-9cf1-9028d094296e/download/ReferenceCCAProfiles20142018.csv)

[Key to the variable names of Chicago Census](https://datahub.cmap.illinois.gov/dataset/1d2dd970-f0a6-4736-96a1-3caeb431f5e4/resource/d23fc5b1-0bb5-4bcc-bf70-688201534833/download/CDSFieldDescriptions202007.pdf)

[Chicago socionomic data](https://data.cityofchicago.org/api/views/kn9c-c2s2/rows.csv)

[Chicago public schools data](https://data.cityofchicago.org/api/views/9xs2-f89t/rows.csv)
```{r}
Chicago_census <- read.csv(url("https://datahub.cmap.illinois.gov/dataset/1d2dd970-f0a6-4736-96a1-3caeb431f5e4/resource/8c4e096e-c90c-4bef-9cf1-9028d094296e/download/ReferenceCCAProfiles20142018.csv"))

Chicago_socionomic <- read.csv(url("https://data.cityofchicago.org/api/views/kn9c-c2s2/rows.csv"))

Chicago_schools <- read.csv(url("https://data.cityofchicago.org/api/views/9xs2-f89t/rows.csv"))
Chicago_census
Chicago_socionomic
Chicago_schools
```

```{r} 
Chicago_combined <- Chicago_socionomic %>%
  rename(GEOG = "COMMUNITY.AREA.NAME") %>%
  left_join(Chicago_census, by = "GEOG")

Chicago_combined
```

```{r}
Chicago_data <- Chicago_combined %>%
  select(c("PER.CAPITA.INCOME", "HARDSHIP.INDEX"))
Chicago_model <- lm(PER.CAPITA.INCOME ~., data = Chicago_data)
summary(Chicago_model)


?geom_smooth

ggplot(data = Chicago_data, mapping = aes(x = HARDSHIP.INDEX, y = PER.CAPITA.INCOME)) +
  geom_point() +
  geom_smooth(method = "loess")
```

# Visualize Categorical-Categorical Relation

**Research Question:** What is the relationship between type of transit (using percentage) and the hardship index of each neighborhood.

**Graph:**
```{r}
Q3 <- Chicago_combined %>%
  select(c("GEOG", "TOT_COMM", "DROVE_AL", "CARPOOL", "TRANSIT", 
           "WALK_BIKE", "COMM_OTHER", "HARDSHIP.INDEX" )) %>%
  transmute(GEOG = GEOG, DROVE = DROVE_AL / TOT_COMM, CARPOOL = CARPOOL / TOT_COMM, 
            TRANSIT = TRANSIT / TOT_COMM, WALK_BIKE = WALK_BIKE/ TOT_COMM, 
            COMM_OTHER = COMM_OTHER / TOT_COMM, HARDSHIP = HARDSHIP.INDEX) %>%
  pivot_longer(c("DROVE", "CARPOOL", "TRANSIT", "WALK_BIKE", "COMM_OTHER"), names_to = "TypeTransit", values_to = "Percentage")

Q3 
ggplot(data = Q3, mapping = aes(x = Percentage, y = HARDSHIP, color =TypeTransit)) +
  geom_point() +
  facet_wrap(TypeTransit~.) + 
  theme(legend.position = "none") 

```

```{r}
Chicago_3 <- lm()
```


**Description (1-sentence):**  

In general, there appears to be a strong positive trend between (neighborhood)hardship index and percentage of those who carpool as transit, a horseshoe/quadratic relationship between those who drove to transit, and the hardship index of neighborhoods, an unrelated trend between transit and hardship index, and an unclear relationship between walking and biking, and the hardship index.

# Visualize Categorical-Continuous Relation

**Research Question:**  What is the relationship between percentage of each ethnicity and most common career industry within each neighborhood.

**Graph:**
```{r}
Q2_data_alt <-  Chicago_combined %>% 
  select(c("TOT_POP", "WHITE", "BLACK", "ASIAN", "OTHER", 
           "RES_NAICS1_TYPE")) %>%
  transmute(WHITE_percent = WHITE / TOT_POP, BLACK_percent = BLACK / TOT_POP, 
            ASIAN_percent = ASIAN / TOT_POP, OTHER_percent = OTHER/ TOT_POP, 
            JOB = RES_NAICS1_TYPE) %>%
  pivot_longer(c("WHITE_percent", "BLACK_percent", "ASIAN_percent", "OTHER_percent"), names_to = "Ethnicity", values_to = "Ethnicity_Percentage")
Q2_data_alt
ggplot(data=Q2_data_alt) + 
  geom_boxplot(aes(x=JOB,y=Ethnicity_Percentage,color=Ethnicity)) +
  facet_wrap(Ethnicity~.) +
  coord_flip()

```


**Description (1-sentence):**  This breakdown of ethnicity percentage vs occupation shows Asian residents of each community occupying more accommodation and food service roles and Black and White residents occupying more health care roles.

# Visualize Continuous-Continuous Relation

**Research Question:**  Within each neighborhood, what is the relationship between the percentage without High School Diplomas and Households below poverty level (grouped by quantiles of income-levels)
**Graph:**
```{r} 
Q1 <- Chicago_combined %>%
  select(c("PERCENT.AGED.25..WITHOUT.HIGH.SCHOOL.DIPLOMA", "PERCENT.HOUSEHOLDS.BELOW.POVERTY", 
           "GEOG", "PER.CAPITA.INCOME")) %>%
  mutate(y = PERCENT.HOUSEHOLDS.BELOW.POVERTY,
         z = PER.CAPITA.INCOME,
         x = PERCENT.AGED.25..WITHOUT.HIGH.SCHOOL.DIPLOMA,
         IncomeRank = ntile(z,5)) %>%
  group_by(IncomeRank)
#Plot

P1 <- ggplot(data = Q1, mapping = aes(x = x,y = y,color = IncomeRank)) +
  geom_point() + 
  xlab("Percent aged 25 without a High School Diploma (%)") + 
  ylab("Percent of Households below Poverty (%)") +
  geom_smooth(method = "loess", mapping = aes(x = x,y = y), se = FALSE) 
P1
```

**Description (1-sentence):**  It shows a highly variable relationship, in which a linear, quadratic, or 3rd degree fits do not best account for the pattern. No clear distinction can be made, but between 0 to 30% of those aged 25 without a Highschool Diploma, there seems to be a weak positive correlation for Percent of Households below Poverty. 
